Positional entropy in childes

### 2018-08-17 Unigram Positional Entropy.Rmd
Unigram entropy for Spanish, German and Japanese, plus conditional entropy for Brown corpus of English

```{r Spanish unigram frequencies}
shiro_utterances <- get_utterances(language = "SPA", corpus="Shiro")

@@

  if(role == "Target_Child") {
    sub_utterances <- utterances %>%
      filter(speaker_role == "Target_Child") %>%
      mutate(length = str_count(gloss, " ") + 1) # uses stems of words, not original words
  } else {
     sub_utterances <- utterances %>%
      filter(speaker_role != "Target_Child") %>%
      mutate(length = str_count(gloss, " ") + 1)
  }

@@

entropies <- map(2:10, ~get_entropies("not child", .x, shiro_utterances)) %>%
  bind_rows()

@@

entropies_child <- map(2:10, ~ get_entropies("Target_Child", .x, shiro_utterances)) %>%
  bind_rows()

@@

ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper, color = role)) +
  facet_wrap(~ length) + 
  geom_pointrange(position = position_dodge(.25)) +
  geom_smooth(se = F) +
  ggtitle("Child and Adult Positional Entropy vs. Word Position (Spanish Shiro corpus)")
```

>>>>

```{r Spanish unigram frequencies}
shiro_utterances <- get_utterances(language = "SPA", corpus="Shiro") 
#Venezuelan children narrative (113 children; half 1st graders half 4th graders; half high SES half low SES)

wagner_utterances <- get_utterances(corpus = "Wagner") #German corpus of children during daily routines (age range from 1,5 to 14,10)

okayama_utterances <- get_utterances(corpus = "Okayama") #mother-child conversations in the Osaka area

@@

  if(role == "Target_Child") {
    sub_utterances <- utterances %>%
      filter(speaker_role == "Target_Child") %>%
      mutate(length = str_count(gloss, " ") + 1) 
  } else {
     sub_utterances <- utterances %>%
      filter(speaker_role != "Target_Child") %>%
      mutate(length = str_count(gloss, " ") + 1)
  }

@@

  tokens %>%
    group_by(word_order, word) %>%
    summarise(n = n()) %>%
    tidyboot(summary_function = function(x) x %>% 
               summarise(entropy = entropy.empirical(n, unit = "log2")),
             statistics_functions = function(x) x %>%
             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
    mutate(role = role, length = sen_length)
}

entropies <- map(2:10, ~get_noneng_entropies("not child", .x, okayama_utterances)) %>%
  bind_rows() %>% filter(word_order <= length)

@@

entropies_child <- map(2:10, ~ get_noneng_entropies("Target_Child", .x, okayama_utterances)) %>%
  bind_rows() %>% filter(word_order <= length)

@@

ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper, color = role)) +
  facet_wrap(~ length) + 
  geom_pointrange(position = position_dodge(.25)) +
  geom_smooth(se = F) +
  ggtitle("Child and Adult Positional Entropy vs. Word Position (Japanese Okayama corpus)")
```

### 2018-08-17 positional_entropy_childes.Rmd
Unigram entropy for Spanish, German and Japanese, plus conditional entropy for Brown corpus of English

```{r setup, include = FALSE}
# load packages
library(knitr)
library(tidyverse)
library(directlabels)
library(childesr) #data
library(SnowballC) #stemmer
library(tidytext)
library(entropy)
library(tidyboot)
library(dplyr)
library(tokenizers)
library(tau)

@@

>>>>

```{r setup, include = FALSE}
# load packages
library(knitr)
library(tidyverse)
library(directlabels)
library(childesr) #data
library(SnowballC) #stemmer
library(tidytext)
library(entropy)
library(tidyboot)
library(dplyr)
library(tokenizers)
library(tau)
library(gtools)

@@

```{r conditional_entropy}
get_counts <- function(role, sen_length, utterances) {

all_conditional_entropy <- function(df) {
  max_len <- max(df$word_order)
  #combs <- combn(1:max_len, 2, simplify = F)
  perms <- permutations(max_len, 2, 1:max_len, repeats.allowed = F) 
  combs <-  setNames(split(perms, seq(nrow(perms))), rownames(perms))

  map(combs, ~conditional_entropy(df, .x[1], .x[2])) %>%
    bind_rows()

conditional_entropy <- function(df, X, Y) { #Y is conditioning variable

  d <- df %>%
    filter(word_order %in% c(X, Y)) %>%
    mutate(word_order = factor(word_order, labels = c("first", "second"))) %>%
    spread(word_order, word)

  ent <- d$first %>% table() %>% as.vector() %>% entropy.empirical(unit="log2")

  mi <- d %>% 
    group_by(first, second) %>%
    summarise(n = n()) %>%
    spread(second, n, fill = 0) %>% 
    ungroup() %>%
    select(-first) %>%
    mi.empirical(., unit = "log2")

  data_frame(length = max(df$word_order), X = X, Y = Y, mi = mi, ent = ent, ce = ent - mi)
}

ce <- map(counts, all_conditional_entropy)

ggplot(mi, aes(x = pos2, y = mi, color = as.factor(pos1))) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_line()

ggplot(ce[[5]], aes(x = X, y = ce, color = as.factor(Y))) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Conditional entropy vs. conditioned random variable (utterances of length 6)")
